---
title: "Homework 2- Correlating like a Data Master"
author: Riyanshi Bohra
format: html
editor: visual
---

# Homework 2-

# Correlating like a Data Master

# Setting up

```{r}
options(digits = 2)

if (!require(pacman))
  install.packages("pacman")

# Installing BiocManageR
if (!require(BiocManager))
  install.packages("BiocManager")

pacman::p_load(
       dlookr,
       formattable,
       GGally,
       ggdist,
       ggpubr,
       ggridges,
       here,
       tidyverse)

# Set global ggplot() theme
theme_set(theme_pubclean(base_size = 12)) 

theme_update(axis.title = element_text(hjust = 1))

theme_update(axis.ticks = element_blank()) 

theme_update(legend.key = element_blank())
```

## Loading the Tornados dataset

```{r}
dataset <- read.csv("/Users/riyanshibohra/Desktop/tornados.csv") 
```

# Diagnosing the data

```{r}
# What does the data look like?
dataset |>
  head() |>
  formattable()
```

# Describing and Visualizing Correlations

## Table of correlations: Numeric variables

```{r}
correlate(dataset) |>
  formattable()
```

## Correlation matrix: Numerical variables

```{r}
dataset |>
plot_correlate()
```

## Visualize Correlations within Groups

```{r}

# Grouping by number of states(ns) affected by this tornado. 1, 2, or 3.
dataset |>
  group_by(ns) |>
  plot_correlate()
```

## Using ggpairs()

```{r}
# Define the Okabe-Ito colors
#OkabeItoColors <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

#dataset |> 
 # dplyr::select(st, len, wid) |>
  #ggpairs(aes(color = st, alpha = 0.5)) +
 # theme(strip.background = element_blank()) + 
  #scale_fill_OkabeIto() +
  #scale_color_OkabeIto()
#The categorical column in this dataset has too many categories which is why this code is giving error saying: Column 'st' has more levels (53) than the threshold (15) allowed. #Possible options to deal with this is:
#1. Increase the cardinality_threshold Parameter. But according to the error message this will produce long processing times
#2. Reduce the Cardinality of 'st' Column. We can't removes categories from the column because that would give us the wrong analysis. 
#3. Convert to Numeric. But this function doesn't work for numeric data
```

#Visualizing relationship based on the Target Variable

##Numerical Target Variables: Numerical predictor variable

```{r}
# Removing NA values
dataset.noNA <- dataset |> 
  drop_na()
```

```{r}
# The target variable
num <- target_by(dataset.noNA, sn)

```

```{r}
# Relating the predictor variable and target variable
num_num <- relate(num, ns)

# Summary of the regression analysis 
summary(num_num)
```

```{r}
# Plotting the linear relationship
plot(num_num)
```

## Numerical Target Variables: Categorical predictor variable

```{r}
# The target variable
num <- target_by(dataset, ns) 
```

```{r}

num$tz <- as.factor(num$tz)

# Relating the predictor variable and the target variable
num_cat <- relate(num, tz)

# Summary of the ANOVA analysis \
summary(num_cat)
```

```{r}
#Plotting the graph
plot(num_cat) + 
  theme(axis.text.x = element_blank())
```

## Categorical Target Variable: Numerical predictor variable

```{r}
# The categorical predictor variable that we want
categ <- target_by(dataset, tz)

# Relating the variable of interest to the numerical target variable
cat_num <- relate(categ, wid)

# Summary of descriptive statistics
summary(cat_num)
```

```{r}
#Plotting

plot(cat_num) 
```
## Categorical Target Variable: Categorical predictor variable
```{r}
# Creating new categorical column
cat_dataset <- dataset |>
  select(wid, tz) |>
  drop_na() |>
  mutate(HighLWP = ifelse(
    wid > (mean(wid + sd(wid))), 
                          "Yes", 
                          "No"))

# New dataset 
cat_dataset |>
  head() |>
  formattable()
```

```{r}
# The categorical predictor variable that we want
categ <- target_by(cat_dataset, HighLWP)

# Relating the variable of interest to the categorical target variable
cat_cat <- relate(categ, tz)

# Summary of the Chi-square test for Independence
summary(cat_cat)
```

```{r}
#Plotting
plot(cat_cat)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
