---
title: "Homework 2- Transforming like a Data Transformer"
author: Riyanshi Bohra
format: html
editor: visual
---

# Homework 2-

# Transforming like a Data Transformer

# Setting up

```{r}
options(digits = 2)

# Installing and loading required packages
if (!require(pacman))
  install.packages("pacman")

pacman::p_load(dlookr, 
               formattable, 
               here, 
               kableExtra, 
               knitr, 
               missRanger, 
               tidyverse) 
```

## Loading the Tornados dataset and adding a categorical group

```{r}
dataset <- read.csv("/Users/riyanshibohra/Desktop/tornados.csv") |>
   # Add a categorical group
  mutate(Year_group = ifelse(yr >= 1950 & yr <= 1975, "1950-1975", 
                            ifelse(yr > 1975 & yr <= 2000, "1976-2000", 
                                   "2001-2022")),
         Year_group = fct_rev(Year_group))
```

# Diagnosing the data

```{r}

# What does the data look like?
dataset |>
  head() |>
  formattable()
```

# Describing the properties of Data

```{r}
dataset |>
  select(slat, slon, elat, elon) |>
  describe() |>
  select(described_variables, skewness) |>
  formattable()
```

# Testing Normality

## Q-Q Plots

```{r}
dataset |>
plot_normality(slat, slon, elat,elon)
```

# Normality within Groups

## Q-Q Plots

```{r}
dataset %>%
  group_by(Year_group) %>%
  select(slat, slon) %>%
  plot_normality()
```

# Transforming the Data

```{r}
#Transforming the width in yards column(wid)
WidMod <- dataset |>
  filter(wid > 0)
```

##Square-root Transformation

```{r}
sqrtWid <- transform(WidMod$wid, method = "sqrt") 

summary(sqrtWid)
```

```{r}
#Plotting the transformed data
sqrtWid |>
  plot()
```

## Logarithmic Transformation

```{r}
Log1Wid <- transform(WidMod$wid, method = "log+1") 

summary(Log1Wid)
```

```{r}
# Plotting the transformed data
Log1Wid |>
  plot()
```

## Inverse Transformation

```{r}
InvWid <- transform(WidMod$wid, method = "1/x") 

summary(InvWid)
```

```{r}
# Plotting the transformed data
InvWid |>
  plot()
```
## Squared Transformation
```{r}
SqrdWid <- transform(WidMod$wid, method = "x^2") 

summary(SqrdWid)
```

```{r}
SqrdWid |>
  plot()
```
## Cubed Transformation
```{r}
CubeWid <- transform(WidMod$wid, method = "x^3") 

summary(CubeWid)
```

```{r}
CubeWid |>
  plot()
```

## Box-cox Transformation

```{r}
BoxCoxWid <- transform(WidMod$wid, method = "Box-Cox") 

summary(BoxCoxWid)
```

```{r}
# Plotting the transformed data
BoxCoxWid |>
  plot()
```

# Producing and HTML Summary
```{r}
#transformation_web_report(dataset)
# Without commenting out this code chunk, the file isn't getting rendered
```

